---
title: "FinalProjectPresi"
author: "Jax Lubkowitz"
format: 
  revealjs:
    embed-resources: true
editor: visual
execute:
  echo: false
  warning: false
---

## Data Introduction
- Patients with Multiple-sclerosis, a chronic neuroinflammatory autoimmune disease, inhibiting the CNS

 - autoimmune disease -> serious implications in pandemic
 
 - What factors impact covid19 in MS?
 
 - Sample: 1141 people with MS
 
 
## Data 
```{r}
library(tidyverse)
library(here)
covid_df <- read_csv(here("data/physionet.org/files/patient-level-data-covid-ms/1.0.0/GDSI_OpenDataset_Final.csv"))

age_df <- covid_df |> select(age_in_cat, covid19_confirmed_case) |> 
  mutate(age_in_fac = as_factor(age_in_cat)) |> 
  mutate(age_in_fac = case_when(
    age_in_fac == 0 ~ "0-18",
    age_in_fac == 1 ~ "19-50",
    age_in_fac == 2 ~ "51-70",
    age_in_fac == 3 ~ "71+")) |> 
  rename(Response = "covid19_confirmed_case") |>
  mutate(Response = if_else(Response == "yes", "Positive", "Negative"))


ggplot(data = age_df, aes(x = age_in_fac, fill = Response)) +
  geom_bar() +
  labs(title = "Number of COVID cases by Age", x = "Age", y = "Total")
```

## With Assumption that Suspected Covid are confirmed cases
```{r}
cofirmed_df <- covid_df |> 
  mutate(had_covid = case_when(
    covid19_diagnosis == "suspected" ~ "Positive",
    covid19_diagnosis == "confirmed" ~ "Positive",
    covid19_diagnosis == "not_suspected" ~ "Negative")) |>
  mutate(covid19_diagnosis = as_factor(covid19_diagnosis)) |> 
  mutate(age_in_fac = as_factor(age_in_cat)) |>
  mutate(covid19_diagnosis = case_when(
    age_in_fac == 0 ~ "0-18",
    age_in_fac == 1 ~ "19-50",
    age_in_fac == 2 ~ "51-70",
    age_in_fac == 3 ~ "71+")) |> 
  rename(Response = "had_covid")


ggplot(data = cofirmed_df, aes(x = age_in_fac, fill = Response)) +
  geom_bar() +
  labs(title = "Number of COVID cases by Age", x = "Age", y = "Total")
```


## ChiSQR of covid Hospotaliztion of MS population with/without Comorbidities
Check conditions?

Ho: no association of covid outcome and Comorbidities
Ha: is related

```{r}
CHI <- covid_df |> 
  mutate(covid19_outcome = as_factor(covid19_outcome_levels_2)) |>
  mutate(covid19_outcome = case_when(
    covid19_outcome == 0 ~ 'Not Hospitlized',
    covid19_outcome == 1 ~ 'Hospitlized',
    covid19_outcome == 2 ~ 'ICU or VF',
    covid19_outcome == 3 ~ 'Death'))

chisq.test(x = CHI$covid19_outcome, y = CHI$has_comorbidities)
```

##  Comorbidities associtation with Hospitlization pt. 1
```{r}
chisq.test(x = CHI$covid19_outcome, y = CHI$com_cardiovascular_disease) #nope
chisq.test(x = CHI$covid19_outcome, y = CHI$com_chronic_kidney_disease) # yup
chisq.test(x = CHI$covid19_outcome, y = CHI$com_chronic_liver_disease) # yup
chisq.test(x = CHI$covid19_outcome, y = CHI$com_diabetes) #yup

```
 
##  Comorbidities associtation with Hospitlization pt. 2
```{r}
chisq.test(x = CHI$covid19_outcome, y = CHI$com_immunodeficiency) # yup
chisq.test(x = CHI$covid19_outcome, y = CHI$com_lung_disease) # nope
chisq.test(x = CHI$covid19_outcome, y = CHI$com_malignancy) #nope
chisq.test(x = CHI$covid19_outcome, y = CHI$com_neurological_neuromuscular) # yup
#chisq.test(x = CHI$covid19_outcome, y = CHI$comorbidities_other) # Not enough data
```

