---
title: "TakeHomeQuiz11"
author: "Jax Lubkowitz"
format: 
  html:
    embed-resources: true
---
Statement of Integrity: “All work presented is my own, and I have followed all rules for collaboration. If I used any AI on this assignment, I have clearly stated both (1) what my search prompt was and (2) what I used from the AI answer.””

Jax Lubkowitz


## 1 Jessie Diggins (fastest US Skier) vs Therese Johaug (fastest skier in world)
The huge Jessie Diggens spike in 2018 is when she won the first ever US gold medal at the Olympics for Nordic Skiing.
```{r}
library(here)
library(lubridate)
library(tidyverse)
skiers <- read_csv(here("data/skiers.csv"), skip = 2, na = c("", "NA", '<1'), col_types = cols(
  `Jessie Diggins: (United States)` = col_double(),
  `Therese Johaug: (United States)` = col_double()))

skiers <- skiers |> rename(JD = `Jessie Diggins: (United States)`, TJ = `Therese Johaug: (United States)`) |> mutate(Month = ymd(Month, truncated = 1)) |> mutate(TJ = if_else(TJ == 'NaN', TJ, 0))

ggplot(data = skiers)+
  geom_line(aes(x = Month, y = JD), color = "red") +
  geom_line(aes(x = Month, y = TJ), color = "blue")
```

## 2 Fortnite Trend by Region
```{r}
fortnite_df <- read_csv(here("data/geoMap.csv"), skip = 1, col_types = cols(
  Region = col_character(), 
  'fortnite: (1/1/04 - 11/15/23)' = col_double())) |> rename(Popularity = 'fortnite: (1/1/04 - 11/15/23)')

state_names <- as_tibble(state.name)
state_regions <- as_tibble(state.region)
state_df <- bind_cols(state_names, state_regions) |>
  rename(state_name = `value...1`, state_region = `value...2`)

#Used an inner join to drop Washington DC
regional_fort_df <- fortnite_df |> inner_join(state_df, by = join_by(Region == state_name))

regional_fort_df |> group_by(state_region) |> summarise(average = mean(Popularity))

```



a question on something similar to one of the `lubridate` google trends class exercises and/or your turn exercises.
a question similar to something you completed in Project 1.
a question similar to something you completed in the second in-class coding quiz (`06-quiz-class.html`)
a question similar to something you completed in Project 2.
